---
title: "R Codes - ICRC"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require("knitr")
opts_knit$set(root.dir = "/Desktop/ICRC")
```

## Set your working directory and/or path the data files
## Create a project from R studio ->File > New Project (rstats.wtf/project-oriented-workflow.html)

##################################
## UW data
## There are N individuals that have a ProRT seq in this file.
####################################

```{r #1) Load the data, and select the columns echo=FALSE, eval = F}

library(tidyverse)

uw.rtseq <- read_csv("~/Dropbox/R01.Seattle/RecordLinkage/ProtRt_Sequence_Dec.2018.csv") 
uw.rt <- uw.rtseq[, c(1,3,4)]%>%
  mutate(SequenceType=ifelse(SequenceType=="PrtRt", "PR/RT"))
#Select PtID, SequenceType(converted "PrtRT" to "PR/RT"), Sequence
view (uw.rt)
dim(uw.rt) #[1] 4745    3

uw.person <- "~/Dropbox/R01.Seattle/RecordLinkage/Demographic_Dec.2018.csv" %>% read_csv
view(uw.person)
dim(uw.person)#[1] 2795   11

uw.person.selected <-
    uw.person %>%
    select(PtID, BirthYear, PresentSex) #uw.pt in the original script

```


```{r #2) Assign Sequence ID based on the Patient ID and Column Number, echo=FALSE, eval=F}

uw.rt_split <- uw.rt %>% split(.$PtID)
SeqIDfunction <-function(uwptIDtibble){
  paste(uwptIDtibble$PtID,row.names(uwptIDtibble), sep="-")
}
uw.rt1<-uw.rt %>%
  mutate(SeqID=uw.rt_split %>% lapply(SeqIDfunction)%>%unlist)
view(uw.rt1)
dim(uw.rt1)#[1] 4745    5
```


```{r #3) Combine demographic and sequence data, rename, and selet the columns, and remove duplicates echo=FALSE, eval=F}

# This data will contain only the data that exist in both datasets ####Same as the original script#####
merged.uw.innerjoin<-inner_join(uw.rt1,uw.person.selected, by="PtID")%>%
  rename(birth.year=BirthYear,sex=PresentSex)%>%
  select("PtID", "Sequence", "birth.year", "sex", "SequenceType","SeqID")%>%
  distinct
dim(merged.uw.innerjoin) #[1] 4745    6 ***Same as the original script

```


###########################################
## PHSKC data 
###########################################

```{r #1) Load the data, and clean and rename the columns echo=FALSE, eval=F} 
library(tidyverse)
ph.rtseq <- read.csv("Downloads/PHSKC_PR_RTSeq.csv",header=T)
view(ph.rtseq)
#Sequence and Seq ID 

meta = readRDS(file="Downloads/sequences_meta.rds") 
view(meta)

meta.selected<-meta%>%
  rename(PtID=newnum, SeqID=seqID)%>%
  select("SeqID","PtID")  #SeqID and Pt ID
view(meta.selected)

```

```{r #2) Combine the meta & phrtseq data  echo=FALSE, eval=F}
###*****[dim] does not match with the original script : [1] 18263   4

##1) keeping only the data present in both datasets #####Original Script ##
ph.rtseq.ptid_inner<- inner_join(meta.selected, ph.rtseq, by ="SeqID")%>%
   rename(Sequence=PrtRT.seq)
view(ph.rtseq.ptid_inner) 
dim(ph.rtseq.ptid_inner) #[1] 18262     3 #It has only the data that have all the Sequences/PtID/SeqID 

##2) Keep the data that have PtID & SeqID 
ph.rtseq.ptid_left<- left_join(meta.selected, ph.rtseq, by ="SeqID")%>%
   rename(Sequence=PrtRT.seq)
view(ph.rtseq.ptid_left) 
dim(ph.rtseq.ptid_left) #[1] 21277     3 #This data have some raws without sequences

```

```{r #3 Combine and clean the data with demograph information echo=FALSE, eval=F}

#Load data and select the columns, and remove the duplicates
person <- readRDS(file="Downloads/person.rds")
ph.pt<- person%>%
  select("newnum","b_yr","sex")%>%
  mutate(b_yr=as.double(b_yr))%>%
  rename(PtID=newnum, birth.year=b_yr)
 #Data with PtID, BirthYear, and Sex
view(ph.pt)

#Combine Sequence data with demographic Information, remove duplicates, and separate the "Sequence type" and "Sequence"

##1) Keep the only data that have no missing information #####Original SCRIPT####
ph.rt1_inner<-inner_join(ph.rtseq.ptid_inner, ph.pt,by="PtID")%>%
  select("PtID","Sequence","birth.year","sex","SeqID")%>%
  separate(SeqID, into = c("SequenceType", "SeqID"), sep = "-")%>%
  distinct 
view(ph.rt1_inner)#This will contain the only unique data having both sequences / Sequence ID 
dim(ph.rt1_inner) #[1] 18262     6

##2) Keep the data with PtID and SeqID (Including those with no Sequence data available)
ph.rt1_left<-left_join(ph.rtseq.ptid_left, ph.pt,by="PtID")%>%
  select("PtID","Sequence","birth.year","sex","SeqID")%>%
  separate(SeqID, into = c("SequenceType", "SeqID"), sep = "-")%>%
  distinct 
view(ph.rt1_left)#This will contain all the unique data that have PtID and SeqID, but some missing Sequence data
dim(ph.rt1_left) #[1] 21277     6

```

#############################
########Start to compare cleaned.merged.uw & ph.rt1 by (birth year, sex, seq)
#############################

```{r #3  Extract the overlapping samples from "cleaned.merged.uw and ph.rt1" by birth year, sex and Sequence echo=FALSE, eval=F)}

# Combine the UW and PHSKC data by extracting the data that match by sex, sequence and birth year, removing duplicates (Keeping only those with all the information, no missing data)
 
joined<-inner_join(merged.uw.innerjoin,ph.rt1_inner,by=c("sex","Sequence","birth.year"))%>%distinct%>%
  rename(uw.PtID=PtID.x, uw.SeqID=SeqID.x,ph.PtID=PtID.y, ph.SeqID=SeqID.y, uw.SequenceType=SequenceType.x, ph.SequenceType=SequenceType.y)%>%
  select(uw.PtID,uw.SeqID,ph.PtID,ph.SeqID,uw.SequenceType,ph.SequenceType,birth.year,sex,Sequence)
  
view(joined)

```


```{r #3 Create a single merged dataset with all the unique sequences found in either data set without duplicates. echo=FALSE, eval=F)}

merged.uw.full <- full_join(uw.rt1,uw.person.seleted,by="PtID")%>%
  rename(birth.year=BirthYear,sex=PresentSex)%>%
  select("PtID", "Sequence", "birth.year", "sex", "SequenceType","SeqID")%>%
  distinct
view(merged.uw.full)

ph.rtseq.ptid_full<- full_join(meta.selected, ph.rtseq, by ="SeqID")%>%
   rename(Sequence=PrtRT.seq)%>%
  distinct%>%
   filter(Sequence!=F)

ph.rt1_full<- full_join(ph.rtseq.ptid_full, ph.pt, by ="PtID")%>%
  select("PtID","Sequence","birth.year","sex","SeqID")%>%
  separate(SeqID, into = c("SequenceType", "SeqID"), sep = "-")%>%
   filter(Sequence!=F)%>%
  distinct 

dim(merged.uw.full) #[1] 4803    6
dim(ph.rt1_full)#[1] 18263     6

full_merged<-full_join(merged.uw.full,ph.rt1_full,by="Sequence")%>%
  distinct%>%
  rename(uw.PtID=PtID.x, uw.SeqID=SeqID.x,ph.PtID=PtID.y, ph.SeqID=SeqID.y, uw.SequenceType=SequenceType.x, ph.SequenceType=SequenceType.y, uw.birth.year=birth.year.x, ph.birth.year=birth.year.y,uw.sex=sex.x, ph.sex=sex.y)

full_merged$Sequence%>%unique%>%length #[1] 1900４unique sequences in this merged dataset
full_merged$Sequence%>%length #[1] 19210 Unique Sequences 

full_unique_seq<-full_merged %>% distinct(Sequence, .keep_all = TRUE)
#Keep unique sequences (remove duplicates)
view(full_unique_seq)
view(full_merged)
dim(full_unique_seq) #[1] 19004     11
 
```

###If they work, then I would like to create functions for combining  datasets, removing duplicates, renaming the columns. So that we can use a very short coding for doing this script. ### 
